---
title: "Railway trimmed"
author: "Elia Gonzato"
date: "2024-12-19"
output: pdf_document
---

# Libraries


```{r, warning=F,message=F, echo=FALSE}
library(pacman)
p_load(metafor,
       meta,
       readxl,
       grid,
       splines,
       gdata,
       dplyr,
       ggplot2,
       forestploter,
       writexl,
       dosresmeta)
```
\
Three datasets are needed to perform analysis, specifically: `meta`, `metaregr` and `*.ihme`.

# Dataset for the analysis

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# trim dataset
mi.trim=mi.ihme %>%
  mutate(studyid=study_id) %>% 
  select(studyid, is_outlier) %>% 
  arrange(studyid)
```

# MI

Analysis for myocardial infarction
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# filter for the outcome of interest
mi=meta %>% 
  filter(disease=='Acute myocardial infarction') %>% 
  arrange(studyid)
# filter other rows
mi=mi[-c(3),]
```
\
Compute fixed and random effects models
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# fixed effects
fixed.mi=metagen(beta,sebeta,data=mi,
                     studlab=paste(author,sep=","),
                     comb.fixed=TRUE,comb.random=FALSE,sm="RR")
# random effects
random.mi=metagen(beta,sebeta,data=mi,
                     studlab=paste(author,sep=","),
                     comb.fixed=FALSE,comb.random=TRUE,
                     method.tau="PM",prediction=TRUE,hakn=TRUE,sm="RR")
```
\
Summary of fixed effects model:
\
```{r, echo=FALSE, warning=FALSE, message=FALSE, results = 'hide'}
summary(fixed.mi)
```
\
Summary of mixed effects model:
\
```{r, echo=FALSE, warning=FALSE, message=FALSE, results = 'hide'}
summary(random.mi)
```


### Forest plot

Forest plot of the chosen model:
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
meta::forest(fixed.mi,
             sortvar=country,
             #fill='lightgrey',
             #layout = "BMJ",
             #type.study='circle',
             psize=rep(0.3,11),
             lty.equi = 2,
             cex = 0.3,
             #circlesize=0.8,
             #rightlabs=c('RR', "95% CI", "Weight"),
             header.line='both',
             weight.study='same',
             col.circle='black',
             #col.lines='black',
             ff.random='bold',
             ff.predict='bold',
             ff.hetstat='bold',
             ff.test.overall='bold',
             ff.random.labels = 'bold',
             col.diamond.random='navy',
             col.predict='red',
             leftcols=c("studlab", "year", "country","metric"),
             leftlabs = c("Study", "Year", "Country","Outcome"))
```


## Meta-regression

Meta-regression to fit exposure-response function with splines
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# filter dataset
miregr= metaregr %>% filter(disease=='Acute myocardial infarction')
# add outlier column
miregr$outlier=mi.trim$is_outlier
# filter
miregr=miregr %>% filter(outlier==0)
# create splines object on the observed noise level exposure
spline_basis=ns(miregr$noiselevel, df = 3)
# compute meta-regression
res_spline=rma(yi = beta, sei = sebeta, 
                  mods = spline_basis, 
                  random = ~ 1 | author,
                  data = miregr, 
               method = "REML")
```
\
With the new object, compute predictions over the interval, to obtain the exposure response function
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# create new level of exposure
new_exposure=seq(min(miregr$noiselevel),
                 max(miregr$noiselevel), length.out = 100)
# fit splines to the new exposure
new_spline_basis=ns(new_exposure, df = 3)
# predict 
pred_spline1=predict(res_spline, newmods = new_spline_basis)
# create df with predictions
dfsplines1=data.frame(pred=pred_spline1$pred,
                     lower=pred_spline1$ci.lb,
                     upper=pred_spline1$ci.ub,
                     exposure=new_exposure)
```
\
Plot exposure response function with points
\
```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width="50%", fig.align="center"}
ggplot(miregr,aes(x=noiselevel,y=beta))+
  geom_point(aes(x=noiselevel,y=beta,size=sebeta),alpha=.5,col="#005c99")+
  #geom_label_repel()+
  guides(size='none')+
  labs(x='Noise Exposure',
       y='Log Relative Risk')+
  scale_x_continuous(breaks=seq(45,80,5))+
  geom_line(data=dfsplines1,aes(x=exposure,y=pred), col="#66a3ff")+
  geom_ribbon(data=dfsplines1,aes(x=exposure,y=pred,ymin = lower, ymax = upper),
              alpha = 0.2, col="#66a3ff",fill="#66a3ff") +  # Add confidence intervals
  theme_bw()+
  geom_hline(yintercept = 0,lty=2,col='black')
```


### Funnel plot

Funnel plot for publication bias assessment:
\
```{r, echo=FALSE, warning=FALSE, message=FALSE, out.width="50%", fig.align="center"}
residuals=residuals(res_spline)
miregr$residuals=residuals
# extract estimate
estimate=0
se=res_spline$se[1]
se.seq=seq(0, max(miregr$sebeta)+0.02, 0.01)
#Compute vectors of the lower-limit and upper limit values for
#the 95% CI region
ll95 = estimate-(1.96*se.seq)
ul95 = estimate+(1.96*se.seq)
#
seq.mean=seq(0,0.25,by=0.1)
#You can ignore it.
dfcenter=data.frame(y=estimate,
                    x=seq(0,max(miregr$sebeta)+0.02,0.01))
dfCI = data.frame(ll95, ul95, se.seq, estimate)
# Define fill colors for contour
ggplot(data=miregr,aes(x=sebeta,y=residuals))+
  geom_point(data=miregr,aes(x=sebeta,y=beta,size=sebeta,alpha=0.7))+
  geom_line(aes(x = se.seq, y = ll95), data = dfCI,col='black',alpha=0.8,lty=2) +
  geom_line(aes(x = se.seq, y = ul95), data = dfCI,col='black',alpha=0.8,lty=2) +
  #geom_line(aes(x = se.seq, y = ll99), data = dfCI,col='red') +
  #geom_line(aes(x = se.seq, y = ul99), data = dfCI,col='red') +
  geom_line(aes(x=x,y=y),col='black',data=,dfcenter,alpha=0.8,lty=2)+
  scale_x_reverse()+
  scale_y_continuous(breaks=seq(-1.25,2,0.25))+
  coord_flip()+
  guides(size='none',
         alpha='none')+
     annotate(geom = "polygon", x = c(max(se.seq), 0, max(se.seq)),
              y = c(min(dfCI$ll95), estimate,max(dfCI$ul95)), fill = "gray", alpha = 0.1 )+
  theme_bw()+
  labs(x='Standard Error',
       y='Residuals')
```
\
Egger's test to test presence of publication bias:
\
```{r}
#Egger's test
egger_test=lm(beta ~ sebeta, data = miregr)
summary(egger_test)
et_pval=summary(egger_test)$coefficients[2, 4]
print(et_pval)
```

# Stroke

## Meta-Analysis

Analysis for stroke
\
```{r, echo=FALSE, warning=FALSE, message=FALSE, results = 'hide'}
keep(meta, metaregr, sure = T)
# trim dataset
stroke.trim=stroke.ihme %>%
  mutate(studyid=study_id) %>% 
  select(studyid, is_outlier) %>% 
  arrange(studyid)
```
\
Filter by disease and rows
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# filter for the outcome of interest
stroke=meta %>% 
  filter(disease=='Stroke') %>% 
  arrange(studyid)
# filter other rows
stroke=stroke[-c(1),]
```
\
Compute fixed and random effects models
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# fixed effects
fixed.stroke=metagen(beta,sebeta,data=stroke,
                     studlab=paste(author,sep=","),
                     comb.fixed=TRUE,comb.random=FALSE,sm="RR")
# random effects
random.stroke=metagen(beta,sebeta,data=stroke,
                     studlab=paste(author,sep=","),
                     comb.fixed=FALSE,comb.random=TRUE,
                     method.tau="PM",prediction=TRUE,hakn=TRUE,sm="RR")
```
\
Summary of fixed effects model:
\
```{r, echo=FALSE, warning=FALSE, message=FALSE, results = 'hide'}
summary(fixed.stroke)
```
\
Summary of mixed effects model:
\
```{r, echo=FALSE, warning=FALSE, message=FALSE, results = 'hide'}
summary(random.stroke)
```


### Forest plot

Forest plot of the chosen model:
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
meta::forest(fixed.stroke,
             sortvar=country,
             #fill='lightgrey',
             #layout = "BMJ",
             #type.study='circle',
             psize=rep(0.3,11),
             lty.equi = 2,
             cex = 0.3,
             #circlesize=0.8,
             #rightlabs=c('RR', "95% CI", "Weight"),
             header.line='both',
             weight.study='same',
             col.circle='black',
             #col.lines='black',
             ff.random='bold',
             ff.predict='bold',
             ff.hetstat='bold',
             ff.test.overall='bold',
             ff.random.labels = 'bold',
             col.diamond.random='navy',
             col.predict='red',
             leftcols=c("studlab", "year", "country","metric"),
             leftlabs = c("Study", "Year", "Country","Outcome"))
```


## Meta-regression

Meta-regression to fit exposure-response function with splines
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# filter dataset
strokeregr = metaregr %>% filter(disease=='Stroke')
# add outlier column
strokeregr$outlier=stroke.trim$is_outlier
# filter
strokeregr=strokeregr %>% filter(outlier==0)
# create splines object on the observed noise level exposure
spline_basis=ns(strokeregr$noiselevel, df = 3)
# compute meta-regression
res_spline=rma(yi = beta, sei = sebeta, 
                  mods = spline_basis, 
                  random = ~ 1 | author,
                  data = strokeregr, 
               method = "REML")
```
\
With the new object, compute predictions over the interval, to obtain the exposure response function
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
new_exposure=seq(min(strokeregr$noiselevel),
                 max(strokeregr$noiselevel), length.out = 100)
# Step 3: Create natural splines for the new exposure values for prediction
new_spline_basis=ns(new_exposure, df = 3)

# Step 4: Predict using the same spline basis for the new exposure values
pred_spline1=predict(res_spline, newmods = new_spline_basis)
# create df
dfsplines1=data.frame(pred=pred_spline1$pred,
                     lower=pred_spline1$ci.lb,
                     upper=pred_spline1$ci.ub,
                     exposure=new_exposure)
```
\
Plot exposure response function with points
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(strokeregr,aes(x=noiselevel,y=beta))+
  geom_point(aes(x=noiselevel,y=beta,size=sebeta),alpha=.5,col="#005c99")+
  #geom_label_repel()+
  guides(size='none')+
  labs(x='Noise Exposure',
       y='Log Relative Risk')+
  scale_x_continuous(breaks=seq(45,80,5))+
  geom_line(data=dfsplines1,aes(x=exposure,y=pred), col="#66a3ff")+
  geom_ribbon(data=dfsplines1,aes(x=exposure,y=pred,ymin = lower, ymax = upper),
              alpha = 0.2, col="#66a3ff",fill="#66a3ff") +  # Add confidence intervals
  theme_bw()+
  geom_hline(yintercept = 0,lty=2,col='black')
```


### Funnel plot

Funnel plot for publication bias assessment:
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
residuals=residuals(res_spline)
strokeregr$residuals=residuals
# extract estimate
estimate=0
se=res_spline$se[1]
se.seq=seq(0, max(strokeregr$sebeta)+0.02, 0.01)
#Compute vectors of the lower-limit and upper limit values for
#the 95% CI region
ll95 = estimate-(1.96*se.seq)
ul95 = estimate+(1.96*se.seq)
#
seq.mean=seq(0,0.25,by=0.1)
#You can ignore it.
dfcenter=data.frame(y=estimate,
                    x=seq(0,max(strokeregr$sebeta)+0.02,0.01))
dfCI = data.frame(ll95, ul95, se.seq, estimate)
# Define fill colors for contour
ggplot(data=strokeregr,aes(x=sebeta,y=residuals))+
  geom_point(data=strokeregr,aes(x=sebeta,y=beta,size=sebeta,alpha=0.7))+
  geom_line(aes(x = se.seq, y = ll95), data = dfCI,col='black',alpha=0.8,lty=2) +
  geom_line(aes(x = se.seq, y = ul95), data = dfCI,col='black',alpha=0.8,lty=2) +
  #geom_line(aes(x = se.seq, y = ll99), data = dfCI,col='red') +
  #geom_line(aes(x = se.seq, y = ul99), data = dfCI,col='red') +
  geom_line(aes(x=x,y=y),col='black',data=,dfcenter,alpha=0.8,lty=2)+
  scale_x_reverse()+
  #scale_y_continuous(breaks=seq(-1.25,2,0.25))+
  coord_flip()+
  guides(size='none',
         alpha='none')+
     annotate(geom = "polygon", x = c(max(se.seq), 0, max(se.seq)),
              y = c(min(dfCI$ll95), estimate,max(dfCI$ul95)), fill = "gray", alpha = 0.1 )+
  theme_bw()+
  labs(x='Standard Error',
       y='Residuals')
```
\
Egger's test to test presence of publication bias:
\
```{r}
#Egger's test
egger_test=lm(beta ~ sebeta, data = strokeregr)
summary(egger_test)
et_pval=summary(egger_test)$coefficients[2, 4]
print(et_pval)
```

# Ischemic stroke

## Meta-Analysis

Analysis for ischemic stroke
\
```{r}
keep(meta, metaregr, sure = T)
# trim dataset
istroke.trim=istroke.ihme %>%
  mutate(studyid=study_id) %>% 
  select(studyid, is_outlier) %>% 
  arrange(studyid)
```
\
Filter by rows and outcome of interest
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# filter for the outcome of interest
istroke=meta %>% 
  filter(disease=='Ischemic stroke') %>% 
  arrange(studyid)
# filter other rows
istroke=istroke[-c(2,3),]
```
\
Compute fixed and random effects models
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# fixed effects
fixed.istroke=metagen(beta,sebeta,data=istroke,
                     studlab=paste(author,sep=","),
                     comb.fixed=TRUE,comb.random=FALSE,sm="RR")
# random effects
random.istroke=metagen(beta,sebeta,data=istroke,
                     studlab=paste(author,sep=","),
                     comb.fixed=FALSE,comb.random=TRUE,
                     method.tau="PM",prediction=TRUE,hakn=TRUE,sm="RR")
```
\
Summary of fixed effects model:
\
```{r, echo=FALSE, warning=FALSE, message=FALSE, results = 'hide'}
summary(fixed.istroke)
```
\
Summary of mixed effects model:
\
```{r, echo=FALSE, warning=FALSE, message=FALSE, results = 'hide'}
summary(random.istroke)
```


### Forest plot

Forest plot of the chosen model:
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
meta::forest(fixed.istroke,
             sortvar=country,
             #fill='lightgrey',
             #layout = "BMJ",
             #type.study='circle',
             psize=rep(0.3,11),
             lty.equi = 2,
             cex = 0.3,
             #circlesize=0.8,
             #rightlabs=c('RR', "95% CI", "Weight"),
             header.line='both',
             weight.study='same',
             col.circle='black',
             #col.lines='black',
             ff.random='bold',
             ff.predict='bold',
             ff.hetstat='bold',
             ff.test.overall='bold',
             ff.random.labels = 'bold',
             col.diamond.random='navy',
             col.predict='red',
             leftcols=c("studlab", "year", "country","metric"),
             leftlabs = c("Study", "Year", "Country","Outcome"))
```

## Meta-regression

Meta-regression to fit exposure-response function with splines
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# filter dataset
istrokeregr = metaregr %>% filter(disease=='Ischemic stroke')
# add outlier column
istrokeregr$outlier=istroke.trim$is_outlier
# filter
istrokeregr=istrokeregr %>% filter(outlier==0)
# create splines object on the observed noise level exposure
spline_basis=ns(istrokeregr$noiselevel, df = 3)
# compute meta-regression
res_spline=rma(yi = beta, sei = sebeta, 
                  mods = spline_basis, 
                  random = ~ 1 | author,
                  data = istrokeregr, 
               method = "REML")
```
\
With the new object, compute predictions over the interval, to obtain the exposure response function
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# create new level of exposure
new_exposure=seq(min(istrokeregr$noiselevel),
                 max(istrokeregr$noiselevel), length.out = 100)
# fit splines to the new exposure
new_spline_basis=ns(new_exposure, df = 3)
# predict 
pred_spline1=predict(res_spline, newmods = new_spline_basis)
# create df
dfsplines1=data.frame(pred=pred_spline1$pred,
                     lower=pred_spline1$ci.lb,
                     upper=pred_spline1$ci.ub,
                     exposure=new_exposure)
```
\
Plot exposure response function with points
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(istrokeregr,aes(x=noiselevel,y=beta))+
  geom_point(aes(x=noiselevel,y=beta,size=sebeta),alpha=.5,col="#005c99")+
  #geom_label_repel()+
  guides(size='none')+
  labs(x='Noise Exposure',
       y='Log Relative Risk')+
  scale_x_continuous(breaks=seq(45,80,5))+
  geom_line(data=dfsplines1,aes(x=exposure,y=pred), col="#66a3ff")+
  geom_ribbon(data=dfsplines1,aes(x=exposure,y=pred,ymin = lower, ymax = upper),
              alpha = 0.2, col="#66a3ff",fill="#66a3ff") +  # Add confidence intervals
  theme_bw()+
  geom_hline(yintercept = 0,lty=2,col='black')
```


### Funnel plot

Funnel plot for publication bias assessment:
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
residuals=residuals(res_spline)
istrokeregr$residuals=residuals
# extract estimate
estimate=0
se=res_spline$se[1]
se.seq=seq(0, max(istrokeregr$sebeta)+0.02, 0.01)
#Compute vectors of the lower-limit and upper limit values for
#the 95% CI region
ll95 = estimate-(1.96*se.seq)
ul95 = estimate+(1.96*se.seq)
#
seq.mean=seq(0,0.25,by=0.1)
#You can ignore it.
dfcenter=data.frame(y=estimate,
                    x=seq(0,max(istrokeregr$sebeta)+0.02,0.01))
dfCI = data.frame(ll95, ul95, se.seq, estimate)
# Define fill colors for contour
ggplot(data=istrokeregr,aes(x=sebeta,y=residuals))+
  geom_point(data=istrokeregr,aes(x=sebeta,y=beta,size=sebeta,alpha=0.7))+
  geom_line(aes(x = se.seq, y = ll95), data = dfCI,col='black',alpha=0.8,lty=2) +
  geom_line(aes(x = se.seq, y = ul95), data = dfCI,col='black',alpha=0.8,lty=2) +
  #geom_line(aes(x = se.seq, y = ll99), data = dfCI,col='red') +
  #geom_line(aes(x = se.seq, y = ul99), data = dfCI,col='red') +
  geom_line(aes(x=x,y=y),col='black',data=,dfcenter,alpha=0.8,lty=2)+
  scale_x_reverse()+
  #scale_y_continuous(breaks=seq(-1.25,2,0.25))+
  coord_flip()+
  guides(size='none',
         alpha='none')+
     annotate(geom = "polygon", x = c(max(se.seq), 0, max(se.seq)),
              y = c(min(dfCI$ll95), estimate,max(dfCI$ul95)), fill = "gray", alpha = 0.1 )+
  theme_bw()+
  labs(x='Standard Error',
       y='Residuals')
```
\
Egger's test to test presence of publication bias:
\
```{r}
#Egger's test
egger_test=lm(beta ~ sebeta, data = istrokeregr)
summary(egger_test)
et_pval=summary(egger_test)$coefficients[2, 4]
print(et_pval)
```

